Illustrating Model Properties

### 2024-04-24 model.qmd
start with prevalence plots

# State Distribution of Midline Model

@@

>>>>

# Prior State Distribution of Midline Model

@@

# Observed vs Predicted Prevalence

To verify that the model does what it is supposed to, we compare how often some selected diagnoses appear in the data and how often the model predicts these diagnoses to be.

```{python}
#| label: fig-model-prevalences
#| fig-cap: Comparison of predicted vs observed prevalences.
from lyscripts.scenario import Scenario
from lyscripts.plot.utils import Histogram, BetaPosterior, draw
from scripts import paths

nrows, ncols = 2, 2

plt.rcParams.update(shared.get_fontsizes())
plt.rcParams.update(shared.get_figsizes(
  nrows=nrows,
  ncols=ncols,
  aspect_ratio=2.,
  width=17,
))

fig, axes = plt.subplots(
  nrows=nrows,
  ncols=ncols,
  sharex="col",
  sharey="row",
)

for i, t_stage in enumerate(["early", "late"]):
  axes[i,0].set_yticks([])
  axes[i,0].set_ylabel(t_stage)
  for j, midext in enumerate([False, True]):
    content = []
    for lnl, color in zip(
      ["II"],
      [COLORS["blue"]],
    ):
      scenario = Scenario(
        t_stages=[t_stage],
        midext=midext,
        diagnosis={"contra": {"max_llh": {lnl: True}}}
      )
      content.append(
        Histogram.from_hdf5(
          filename=paths.model_dir / "full" / "prevalences_overall.hdf5",
          dataname=scenario.md5_hash("prevalences"),
          color=color,
          label=f"{t_stage = }, {midext = }"
        )
      )
      content.append(
        BetaPosterior.from_hdf5(
          filename=paths.model_dir / "full" / "prevalences_overall.hdf5",
          dataname=scenario.md5_hash("prevalences"),
          color=color,
        )
      )
    draw(
      axes=axes[i,j],
      contents=content,
      xlims=(0., 20.)
    )
    axes[i,j].legend()
